Estimating the contribution of selection to temporal allele frequency change.

### 2019-09-09 manuscript/manuscript.tex
edited
edited disco

\title{}

@@

Presently, there are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact selection has on genome-wide variation
over short timescales in sexual populations remarkably difficult. This is
because the effect of selection on a polygenic trait (such as fitness) is
distributed across loci proportional to their effect sizes, which can lead to
subtle allele frequency shifts \vb{on standing variation} that are difficult to
distinguish from background levels of genetic drift and sampling variance.
However, increasingly genomic experimental evolution studies with multiple
timepoints, and in some cases multiple replicate populations, are being used to
detect selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate
modes of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.
In addition these temporal-genomic studies have begun in wild populations,
\vb{with the goal of finding variants that exhibit frequency changes consistent
with fluctuating selection} \parencite{Bergland2014-ij,Machado2018-cs}. \vb{In
a previous paper, we proposed that one useful signal for understanding the
genome-wide impact of polygenic linked selection detectable from temporal
genomic studies} is the temporal autocovariance in allele frequency changes
\parencite{Buffalo2019-io}.  These covariances are directly estimable from
temporal genomic data and are created when the loci that underly heritable
fitness variation perturb linked neutral alleles; \vb{in contrast}, when
genetic drift acts alone in a closed population, these covariances are zero in
expectation. \vb{Mathematically, temporal covariances occur because it is
  natural to decompose the total variance in allele frequency change across a
  set of time intervals into the variances and covariances in allele frequency
  change among time intervals.  Furthermore, biologically, these covariances
  reflect the extent to which allele frequency changes in one generation
predict changes in another due to a shared fitness background and shared
selection pressures.}

>>>>

\title{Estimating the contribution of selection to temporal allele frequency change. }

@@

Presently, there are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact selection has on genome-wide variation
over short timescales in sexual populations remarkably difficult. This is
because the effect of selection on a polygenic trait (such as fitness) is
distributed across loci proportional to their effect sizes, which can lead to
subtle allele frequency shifts \vb{on standing variation} that are difficult to
distinguish from background levels of genetic drift and sampling variance.
However, increasingly genomic experimental evolution studies with multiple
timepoints, and in some cases multiple replicate populations, are being used to
detect selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate
modes of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.
In addition these temporal-genomic studies have begun in wild populations,
\gc{some} with the goal of finding variants that exhibit frequency changes consistent
with fluctuating selection \parencite{Bergland2014-ij,Machado2018-cs}. \vb{In
a previous paper, we proposed that one useful signal for understanding the
genome-wide impact of polygenic linked selection detectable from temporal
genomic studies} is the temporal autocovariance in allele frequency changes
\parencite{Buffalo2019-io}.  These covariances are directly estimable from
temporal genomic data and are created when the loci that underly heritable
fitness variation perturb linked neutral alleles; \vb{in contrast}, when
genetic drift acts alone in a closed population, these covariances are zero in
expectation. Mathematically, temporal covariances \gc{are useful} because it is
  natural to decompose the total variance in allele frequency change across a
  set of time intervals into the variances and covariances in allele frequency
  change among time intervals.  Furthermore, biologically, these covariances
  reflect the extent to which allele frequency changes in one generation
predict changes in another due to a shared fitness background and shared
selection pressures.
